---
title: "Step 3: Constructing Network Objects"
author: "Joey"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: TRUE
---

# Creating graph objects
If you have been following in the project narrative, we have successfully aggregated our data and created edgelist dataframes. In order to make visualizations using the igraph package, we need to first create graph objects.

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load necessary packages
library(readxl)
library(tidyverse)
library(openxlsx)
library(igraph)
library(here)

# Set Working Directory
setwd("C:/Users/heblj/OneDrive/Career/Professional_Portfolio/Network_Analysis/MigrationAmericas_NetworkAnalysis/Immigration_Networks")


```


## Creating the graph objects
Next, we need to create graph objects with the igraph package by using graph_from_edgelist() function
```{r Making graph objects, include=TRUE}
### Matrices ----------------------------------------------------------------
# Creating Network Objects ------------------------------------------------

  #1990
  g_1990 <- select(el_1990, Origin, Destination) %>% # select only the first two columns for graphing purposes
    as.matrix() %>% # need to convert dataframe to a matrix
    graph_from_edgelist(directed = T) # creates the network graph object

  
  
  #2000
  g_2000 <- select(el_2000, Origin, Destination) %>% # select only the first two columns for graphing purposes
    as.matrix() %>% # need to convert dataframe to a matrix
    graph_from_edgelist(directed = T) # creates the network graph object
  
  
  
  #2010
  g_2010 <- select(el_2010, Origin, Destination) %>% # select only the first two columns for graphing purposes
    as.matrix() %>% # need to convert dataframe to a matrix
    graph_from_edgelist(directed = T) # creates the network graph object
  
  
  #2020
  g_2020 <- select(el_2020, Origin, Destination) %>% # select only the first two columns for graphing purposes
    as.matrix() %>% # need to convert dataframe to a matrix
    graph_from_edgelist(directed = T) # creates the network graph object

```

## Updating Country Codes
Given that our data included the full country names, which will make reading our network visualizations extremely challenging, we need to simply country names into the internationally recognized three letter codes. Note that we are doing this not by changing the names of the vertices in the graph objects, but rather by including an additional vertex attribute called "code".

```{r Re-Coding vertex country labels, include=TRUE}
# Updating Vertex Labels --------------------------------------------------
    
  # Update vertex names according to Official ISO 3166-1 Alpha-3 codes
  country_codes <- list(
    "Argentina"      = "ARG",
    "Bolivia"        = "BOL",
    "Brazil"         = "BRA",
    "Canada"         = "CAN",
    "Chile"          = "CHL",
    "Colombia"       = "COL",
    "Costa Rica"     = "CRI",
    "Ecuador"        = "ECU",
    "El Salvador"    = "SLV",
    "Guatemala"      = "GTM",
    "Honduras"       = "HND",
    "Mexico"         = "MEX",
    "Nicaragua"      = "NIC",
    "Panama"         = "PAN",
    "Paraguay"       = "PRY",
    "Peru"           = "PER",
    "United States"  = "USA",
    "Uruguay"        = "URY",
    "Venezuela"      = "VEN"
  )

  # Convert country list to a named character vector
  codes <- unlist(country_codes)
  
  # Add country code to each graph object
  #1990
    V(g_1990)$code <- codes[V(g_1990)$name]
  
  #2000
    V(g_2000)$code <- codes[V(g_2000)$name]
    
  #2010
    V(g_2010)$code <- codes[V(g_2010)$name]
    
  #2020
    V(g_2020)$code <- codes[V(g_2020)$name]
```

## Adding Additional Vertex Attributes
In the following code chunks we will add additional vertex attributes. These will be related to network analysis metrics that are useful for understanding the dynamics of a network. 

### Degree
Degree represents the number of edges that a vertex possesses. These edges can be qualified as "In", that is edges coming toward the vertex (in the case of migration, a net migration toward the country), or "Out", that is edges going away from the vertex (net migration out of the country toward the country at the other end of the edge). The code below reflects the addition of these attributes to the vertices.

```{r Adding Degree Measures, include=TRUE}
# Adding Degree-related Attributes ------------------------------------------------

    # Degree Total ------------------------------------------------------------
    
    # Combine graphis into list object for efficiency
    graphs_td <- list(g_1990, g_2000, g_2010, g_2020)
    
    # Calculate degree-total scores for each graph
    degree_scores_td <- lapply(graphs_td, degree)
    
    
    V(g_1990)$DegT <- degree_scores_td[[1]]
    V(g_2000)$DegT <- degree_scores_td[[2]]
    V(g_2010)$DegT <- degree_scores_td[[3]]
    V(g_2020)$DegT <- degree_scores_td[[4]]
    
    
    # Degree IN ------------------------------------------------------------
    
    # Calculate degree-in scores for each graph
    degree_scores_di <- lapply(graphs_td, degree, mode = "in")
    
    
    V(g_1990)$DegI <- degree_scores_di[[1]]
    V(g_2000)$DegI <- degree_scores_di[[2]]
    V(g_2010)$DegI <- degree_scores_di[[3]]
    V(g_2020)$DegI <- degree_scores_di[[4]]
    
    
    # Degree OUT ------------------------------------------------------------
    
    # Calculate degree-out scores for each graph
    degree_scores_do <- lapply(graphs_td, degree, mode = "out")
    
    
    V(g_1990)$DegO <- degree_scores_do[[1]]
    V(g_2000)$DegO <- degree_scores_do[[2]]
    V(g_2010)$DegO <- degree_scores_do[[3]]
    V(g_2020)$DegO <- degree_scores_do[[4]]
```

### Centrality Measure: Betweeness
The next attribute added to the vertices is betweeness. Betweeness is a measure that tells how often a vertex sits on the shortest path(s) between other vertices. In other words, it measures how much a vertex acts as an essential bridge between other vertices.

```{r Adding Betweeness measure, include=TRUE}
# Adding Centralization Measures ----
  # Betweeness -------------------------------------------------------
  
  #1990
  V(g_1990)$betweenness <- betweenness(
    g_1990,
    directed = T,
    weights = NULL
  )
  
  #1990
  V(g_2000)$betweenness <- betweenness(
    g_2000,
    directed = T,
    weights = NULL
  )
  
  #2010
  V(g_2010)$betweenness <- betweenness(
    g_2010,
    directed = T,
    weights = NULL
  )
  
  #2020
  V(g_2020)$betweenness <- betweenness(
    g_2020,
    directed = T,
    weights = NULL
  )

```

### GDP-per-Capita Hypothesis
To assess (at this point only qualitatively via visualizations) the extent to which GDP-per-capita might explain
migration patterns, we will next add GDP-per-capita as an additional attribute to the vertices. The underlying hypothesis is that countries with lower GDP-per-capita scores will experience _greater migration_, that is, greater nubmer of outflows or edges that are directed away from the low-GDP-per-capita country to others with greater GDP-per-capita scores.

Given that the periods assessed in this analysis invovle migration at 10yr increments (i.e. 1990, 2000, 2010, 2020), GDP was calculated by taking the average GDP-per-capita over the previous 10years (e.g. for 1990, GDP-per-capita was calulated by taking the mean of GDP-per-capita for each year from 1980 to 1990).

```{r Adding GDP-per-capita, include=TRUE}
### Adding GDP-per-Cap to each Vertex per year ----
    
    # Load the GDP dataframe
    gdpPerCap <- read.xlsx("IMF_GDPperCap_PPP.xlsx", sheet = "10yr Avgs")
    
    # 1990
    idx <- match(V(g_1990)$name, gdpPerCap$COUNTRY)
    V(g_1990)$GDPperCap <- gdpPerCap$`1990`[idx]
    
    # 2000
    idx <- match(V(g_2000)$name, gdpPerCap$COUNTRY)
    V(g_2000)$GDPperCap <- gdpPerCap$`2000`[idx]
    
    # 2010
    idx <- match(V(g_2010)$name, gdpPerCap$COUNTRY)
    V(g_2010)$GDPperCap <- gdpPerCap$`2010`[idx]
    
    # 2020
    idx <- match(V(g_2020)$name, gdpPerCap$COUNTRY)
    V(g_2020)$GDPperCap <- gdpPerCap$`2020`[idx]
```

### Adding attributes to the Edges - Need to start with edglists

If we want to add attributes to the edges in our network visualizations, such as the number of migrants an edge represents (either in absolute or relative terms), we need to load the edgelists and then add the information they contain (specifically aroudn number of migrants) to our network objects. 

```{r Load Edgelists, include=TRUE}
# Edgelists ---------------------------------------------------------------

  # 1990
   el_1990 <- read.csv("Edgelists/EdgeList_1990.csv") # load the csv edgelist
  
  # 2000
   el_2000 <- read.csv("Edgelists/EdgeList_2000.csv") # load the csv edgelist
  
  # 2010
   el_2010 <- read.csv("Edgelists/EdgeList_2010.csv") # load the csv edgelist
  
  # 2020
   el_2020 <- read.csv("Edgelists/EdgeList_2020.csv") # load the csv edgelist
 
  
   
   ### migration flow "weight(s)" ----
    
    E(g_1990)$MigrationFlowWeight <- el_1990$Immigration_absolute
    E(g_1990)$MigrationFlowWeight_PropOrg <- el_1990$per1000_org
    E(g_1990)$MigrationFlowWeight_PropDest <- el_1990$per1000_dest
    
    E(g_2000)$MigrationFlowWeight <- el_2000$Immigration_absolute
    E(g_2000)$MigrationFlowWeight_PropOrg <- el_2000$per1000_org
    E(g_2000)$MigrationFlowWeight_PropDest <- el_2000$per1000_dest
    
    E(g_2010)$MigrationFlowWeight <- el_2010$Immigration_absolute
    E(g_2010)$MigrationFlowWeight_PropOrg <- el_2010$per1000_org
    E(g_2010)$MigrationFlowWeight_PropDest <- el_2010$per1000_dest
    
    E(g_2020)$MigrationFlowWeight <- el_2020$Immigration_absolute
    E(g_2020)$MigrationFlowWeight_PropOrg <- el_2020$per1000_org
    E(g_2020)$MigrationFlowWeight_PropDest <- el_2020$per1000_dest
    
    
```

### On to Visualizations
Now that we have consolidated and updates our igraph objects, we are ready to begin making network visualizations. It was decided that this would be best demonstrated in and R pub, which can be seen here - [Step 4: Visualizing Networks](https://rpubs.com/jtheblj/Step4_VisualizingNetworks)